Bandwidth-Aware Reconfigurable Memory Design with Hybrid Memory Technologies

### 2011-04-06 paper.tex

\subsection{Hardware Configuration}

@@

\noindent{}{\bf{Capacity of each level.}} The maximum availabe capacity of each
level is determined by the cross point of two memory technologies. In our case,
the capacities of SRAM/L2, STT-RAM/L2, and RRAM/L4 are 2MB, 512MB, and 1GB
respectively. At the system initialization, the capacity of each level is
defined to be the maximum available capacity of the level. According to the
reqirement of different applications, the capacities can be reconfigured at
run-time to provide best performance. The memory space of each level can also be
reconfigured to form an internal level as shown in Figure~\ref{fig:hw-config}.
The hardware and software supports for reconfiguration will be presented in the
following sections.\vspace{5pt}

@@

The major system parameter to be reconfigured is the capacity of each memory
level. At each time interval, we determine the upper bound of the capacity of
memory level-i ($s_{iu}$) by mapping the demand bandwidth (DBW) of a specific
application to the system's provided bandwidth curve, i.e.,
$s_{iu}=f^{-1}(\textrm{DBW})$ where $f(x)$ represents the provided bandwidth
curve of the on-chip memory system stored in the look-up-table in a reserved
non-volatile memory space. The demand bandwidth for each memory level is
generated using the prediction engine presented in section~\ref{sec:predict}. In
previous sections, we show that the memory with smaller capacity provides higher
bandwidth. Theoritically, the higher bandwidth provided by the memory system
leads to higher performance (in terms of throughput) and power consumption of
throughput computing applications~\cite{throughput-vs-bandwidth}. Consequently,
we define a lower bound of memory capacity as
$s_{il}=f^{-1}(\textrm{DBW}*(1+\sigma))$, where $\sigma$ is a pre-defined
threshold to make sure possible provided bandwidth increase without much power
overhead. The capacity of memory level-i ($s_i$) is therefore selected to
satisfy $s_{il}\le s_i\le s_{iu}$.

Figure~\ref{fig:reconfig} describes the hardware support for reconfiguration.

reconfiguration support: how reconfigurable (Li -> Li+1 (Li+0.5); hardware
modification to support reconfiguration.

reconfigure the unused capacity:

- if (> si): becomes si+1

- else: unused (or becomes lut)

>>>>

\subsection{Overview}

@@

\noindent{}{\bf{Capacity of each level.}} The maximum availabe capacity of each
level is determined by the cross point of two memory technologies. In our case,
the capacities of SRAM/L2, STT-RAM/L2, and RRAM/L4 are 2MB, 512MB, and 1GB
respectively. Each level of memory if further configured to be multiple banks
the same way as the conventional cache design. At the system initialization, the
capacity of each level is defined to be the maximum available capacity of the
level. According to the reqirement of different applications, the capacities can
be reconfigured at run-time to provide best performance. The memory space of
each level can be reconfigured to form an internal level as shown in
Figure~\ref{fig:hw-config}. The hardware and software supports for
reconfiguration will be presented in the following sections.\vspace{5pt}

@@

The major system parameter to be reconfigured is the capacity of each memory
level. At each time interval, we determine the upper bound of the capacity of
memory level-i ($s_{iu}$) by mapping the demand bandwidth (DBW) of a specific
application to the system's provided bandwidth curve, i.e.,
$s_{iu}=f^{-1}(\textrm{DBW})$ where $f(x)$ represents the provided bandwidth
curve of the on-chip memory system stored in the look-up-table in a reserved
non-volatile memory space. The demand bandwidth for each memory level is
generated using the prediction engine presented in section~\ref{sec:predict}. In
previous sections, we show that the memory with smaller capacity provides higher
bandwidth. Theoritically, the higher bandwidth provided by the memory system
leads to higher performance (in terms of throughput) and power consumption of
throughput computing applications~\cite{throughput-vs-bandwidth}. Consequently,
we define a lower bound of memory capacity as
$s_{il}=f^{-1}(\textrm{DBW}*(1+\sigma))$, where $\sigma$ is a pre-defined
threshold to make sure possible provided bandwidth increase without much power
overhead. The capacity of memory level-i ($s_i$) is therefore selected to
satisfy $s_{il}\le s_i\le s_{iu}$. Furthermore, the rest of the capacity can be
reconfigured to become level-i.5 if necessary. We describe the reconfiguration
algorithm as follows.

<pseudocode>

Our primary reconfigurable design exploits set associativity in conventional
cache organizations. The desirability of such configuration is to exploit the
division of ways already present in a conventional cache organization.
Figure~\ref{fig:reconfig} depicts the block diagram for the organization of our
design. An n-way set associative cache consists of $n$ data and tag arrays. We
divide the each level of on-chip memory into partitions at the granularity of
the $k$-ways, where $k$ is determined by the available capacity range of a
specific memory technology on the provided bandwidth curve. <100+ ways?
experiment section> Reconfiguration will not affect the bits of the address
fields that are used as tag, index, and block offset bits. Modifications to the
conventional cache architecture include:

\begin{itemize}
\item \textbf{Memory status vector} At each memory level, the corrent
configuration is stored in a memory status vector.
\item \textbf{Input and output paths} The input and output data paths are
duplicated. 
\item \textbf{Additional multiplexors} Futhermore, additional wiring and
multiplexors at address decoders and tag comparators are also required to
support the reconfiguration.
\end{itemize}
